### 2022-07-11 script/ms_report.Rmd
Add rules for variant calling with HaplotypeCaller, and report generation

---
title: "MsGenova Project: Data analysis on variant calling from Illumina data"
author: "Gabrièle Adam"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
    html_document:
        highlight: tango
        number_sections: no
        theme: default
        toc: yes
        toc_depth: 3
        toc_float:
            collapsed: no
            smooth_scroll: yes
params:
    result_dir: result_dir
    rmd: "report.Rmd"
    sample: sample
    DP.min: DP.min
    AR.min: AR.min
    region: region
    multiqc: multiqc
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_library, echo=FALSE, warning=FALSE, error=FALSE}
library(tidyverse)
library(DT)
library(ggplot2)
```

```{r define_params, echo=FALSE, warning=FALSE, error=FALSE}
DP.min<-params$DP.min
AR.min<-params$AR.min
workspace<-params$result_dir
region<-params$region
multiqc<-params$multiqc
print(region)
```

# Source
<a download="report.Rmd" href="`r base64enc::dataURI(file = params$rmd, mime = 'text/rmd', encoding = 'base64')`">R Markdown source file (to produce this document)</a>
<a download="MultiQC report" href="`r base64enc::dataURI(file = params$multiqc, mime = 'text/rmd', encoding = 'base64')`">MultiQC report in HTML formata</a>
<a download="Report_snpeff.html" href="`r base64enc::dataURI(file = params$multiqc, mime = 'text/rmd', encoding = 'base64')`">HTML report obtained after snpeff on GenomicsVCF data</a>