### 2022-08-21 script/ms_report.Rmd
Add pindel rules, tdnascan rules, and beginning report

---
title: "MSGenova (Clean version) Rmakrdown Report"
author: "Gabrièle Adam"
date: "2022-08-16"
output:
    html_document:
        highlight: tango
        number_sections: no
        theme: default
        toc: yes
        toc_depth: 3
        toc_float:
            collapsed: no
            smooth_scroll: yes
params:
  DP.min: DP.min
  AR.min: AR.min
  result_dir: result_dir
  rmd: "report.Rmd"
  sample: sample
  trimmo: trimmo
  multiqc: multiqc
  vector: vector
  flagstat: flagstat
  haplotype_caller: haplotype_caller_vcf
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_library, echo=FALSE, warning=FALSE, error=FALSE}
library(tidyverse)
library(DT)
library(ggplot2)
library(plyr)
library(dplyr)
library(shiny)
vector_calling<-params$vector
run_chromomap<-FALSE
DP.min<-params$DP.min
AR.min<-params$AR.min
workspace<-params$result_dir
sample_sheet<-params$sample
trimmo<- params$trimmo
flagstat<- params$flagstat
haplotype_caller<- params$haplotype_caller
setwd(workspace)
```

```{r define_function}
search_identical_insertion<-function(final_table,threashold){
  list_index<-c()
  # we are gonna filter by vector, then order table bu chr, position, and sample. Next compute the distance between 2 lines (0 then che change).Next use a threshold and filter this table
  new_tab<-as_tibble(final_table) %>% dplyr::arrange(vector,Chr,Breakpoint,sample)
  for (i in 1:nrow(new_tab)){
    if(i ==1){
      #case for line =1
      next
    }
    if((new_tab[i,1] == new_tab[i-1,1]) & ((new_tab[i,2] - new_tab[i-1,2]) <= threashold )){
      list_index <- append(list_index,c(i-1,i))
    }
  }
  # now get line in final table
  identical_insertion<-new_tab[unique(list_index),]
  return(identical_insertion)
}

freq_length_bewteen_insertion<-function(final_table){
  # Must be done by sample
  length_between<-c()
  for (v in unique(final_table$sample)){
    for (spl in unique(final_table$sample)){
  new_tab<-as_tibble(final_table)  %>% dplyr::filter(sample==spl & vector==v) %>% dplyr::arrange(Chr,Breakpoint)
  if(nrow(new_tab) != 0){
  for (i in 1:nrow(new_tab)){
    if(i ==1){
      #case for line =1
      next
    }
   if((new_tab[i,1] == new_tab[i-1,1])){
    length_between <-append(length_between,as.integer(new_tab[i,2] - new_tab[i-1,2]))
   }
  }
  }
    }
  }
  return(length_between)
}

read_file<-function(pathway){
#sample<-list.files(getwd())
#vector<-list.files(list.files(getwd())[1])
sample<-list.files(pathway)
vector<-list.files(paste0(pathway,"/",sample[1]))
i=0
for (spl in sample){
  for (v in vector){
    table<-read.table(paste0(pathway,"/",spl,"/",v,"/5.",v,"_insertion.annotated.bed"),header = FALSE,col.names = c("Chr","Breakpoint","SuppRead","TDNA_info","Orientation","Freq","Annotation"))
    if(! nrow(table)==0){
      #add sample and vector columns
      table["sample"]<-spl
      table["vector"]<- v
      if(i==0){
        final_table<-table
        i=i+1
      }else{
        final_table<-rbind(table,final_table)
      }
      i=i+1
    }
  }
}
return(final_table)
}
```

`r ifelse(vector_calling,"# Study Plasmidic insertion","")`

```{r define_variable}
#sample<-list.files(getwd())
#vector<-list.files(list.files(getwd())[1])
pathway=paste0(workspace,"/05_tdnascan/")
sample<-list.files(pathway)
vector<-list.files(paste0(pathway,"/",sample[1]))
# read sample file
sample_sheet<-read.table(sample_sheet,sep="\t",header=TRUE)
```

```{r read_input}
final_tab<-read_file(pathway)
DT::datatable(final_tab)
final_tab2<- final_tab %>% separate(col=TDNA_info,sep=",",into=c("tdna_st","tdna_end"))  %>% mutate(tdna_st=str_replace(tdna_st,"tdna_st:","")) %>% mutate(tdna_end=str_replace(tdna_end,"tdna_end:","")) %>% mutate(tdna_st=str_replace(tdna_st,"-","0")) %>% mutate(tdna_end=str_replace(tdna_end,"-","0")) %>% mutate(tdna_st=as.numeric(as.character(tdna_st))) %>% mutate(tdna_end=as.numeric(as.character(tdna_end))) %>% mutate(length_break=abs(tdna_end - tdna_st)) %>% mutate(breakpoint_end=Breakpoint+length_break) %>% mutate(name=paste(Chr,Breakpoint,vector,sample,sep="-"))  %>% select(name,Chr,Breakpoint,breakpoint_end,vector,sample)
DT::datatable(final_tab2)
```

## By vector, how many insertions are there by chromosome ?{.tabset .tabset-fade}

```{r barplot_snp, results='asis'}
#For one vector, look if it insert at the same position for several sample with a threshold for space between position
for (v in vector){
  sub_table<-filter(final_tab,vector==v) %>% group_by(Chr,sample) %>% dplyr::count(name="nbr_inster_by_chr")
  #datatable(sub_table)
  #count_tab<- count(sub_table,Chr)
  if(nrow(sub_table)> 0){
    cat("###",v,"\n")
  g<- ggplot(sub_table,aes(x=Chr,y=nbr_inster_by_chr,fill=sample))+geom_bar(stat="identity")+xlab("Numbre of insertion by chromosome")+ylab("Chromosome")+ggtitle(paste0("Number of in sertion of vector ",v," by chromosome"))+theme_bw()
  print(g)
  }
  cat('\n\n')
}
```

## What are the size between the insertion

```{r length_between_insertion}
list_of_size<- freq_length_bewteen_insertion(final_tab)
#size<-as.data.frame(list_of_size,rep("insertion",times=length(list_of_size)))
#ggplot(size,aes(x=list_of_size))+geom_density()+xlim(min(list_of_size),max(list_of_size))
size<-as.data.frame(count(list_of_size))
ggplot(size,aes(x=x,y=freq))+geom_point()+ylim(0,(max(size$freq)+1))
ggplot(size,aes(x=factor(x)))+geom_bar(stat = "count")
```

## Are there the same vector insertion for several sample - New version ?
```{r essai_shiny}
#list_size<-freq_length_bewteen_insertion(final_tab)
#cts<-count(list_size)
#print(cts)
#final_tab<-read_file("/home/gabriele.adam/Documents/Msgenova/sortie_msgenova/05_tdnascan/data")
essai<-search_identical_insertion(final_tab,5000)
DT::datatable(essai)
```

`r ifelse(vector_calling,"# how many samples have insertion of which vector and which do not have any - version barplot?","")`

```{r plot_by_chr_tuto, eval=TRUE, echo = TRUE}
#Establish list of sample without any insertion
no_insertion<-c()
#gff3<-"~/Documents/Msgenova/ref_arabido/Ath-Col0-tair10-WG/Arabidopsis_thaliana.TAIR10.41.gff3"
# read gff3 and extract size of chromosome ?? - > Jump to final table
##sequence-region   1 1 30427671
##sequence-region   2 1 19698289
##sequence-region   3 1 23459830
##sequence-region   4 1 18585056
##sequence-region   5 1 26975502
##sequence-region   Mt 1 366924
##sequence-region   Pt 1 154478
table_size_chr<-data.frame(c(paste0("Chr",c(1:5,"Mt","Pt"))),c(30427671,19698289,23459830,18585056,26975502,366924,154478))
colnames(table_size_chr)<-c("Chr","Size")
for (spl in unique(final_tab$sample)){
  sub_table<-dplyr::filter(final_tab,sample==spl)
  if(nrow(sub_table)==0){
    print("yes")
    no_insertion<-append(no_insertion,spl)
  }else{
 bp<- graphics::barplot(table_size_chr$Size,border=NA,names.arg = table_size_chr$Chr,main=paste0("Barplot showing plasmidic insertion on sample ",spl),horiz=TRUE)
 rownames(bp)<-table_size_chr$Chr
    with(sub_table,segments(sub_table$Breakpoint,bp[sub_table$Chr,]-0.5,sub_table$Breakpoint,bp[sub_table$Chr,]+0.5),bp[sub_table$Chr,]+0.5,col=sub_table$vector)
  #print(bp)
  }
}
print("List of vector without any plasmid insertion")
print(no_insertion)

```

`r ifelse(run_chromomap,"#how many samples have insertion of which vector and which do not have any - version chromoMap","")`

```{r chromomap , eval =run_chromomap, echo=run_chromomap}
library(chromoMap)
path<-"/home/gabriele.adam/Documents/Msgenova/sortie_msgenova/chromomap/"
chr<-paste0(path,"chr_size")
annotation<-paste0(path,"annotation.tsv")
for (spl in unique(final_tab$sample)){
  sub_table<-dplyr::filter(final_tab,sample==spl)
  if(nrow(sub_table)==0){
    print("yes")
    no_insertion<-append(no_insertion,spl)
  }else{
 bp<- barplot(table_size_chr$Size,border=NA,names.arg = table_size_chr$Chr,main=paste0("Barplot showing plasmidic insertion on sample ",spl))
 rownames(bp)<-table_size_chr$Chr
  with(sub_table,segments(bp[sub_table$Chr,]-0.5,sub_table$Breakpoint,bp[sub_table$Chr,]+0.5),sub_table$Breakpoint,col=sub_table$vector)
  }}
chromoMap(chr,annotation)
sub_table<-dplyr::filter(final_tab,sample==spl) %>% separate(col=TDNA_info,sep=",",into=c("tdna_st","tdna_end"))  %>% mutate(tdna_st=str_replace(tdna_st,"tdna_st:","")) %>% mutate(tdna_end=str_replace(tdna_end,"tdna_end:","")) %>% mutate(tdna_st=str_replace(tdna_st,"-","0")) %>% mutate(tdna_end=str_replace(tdna_end,"-","0")) %>% mutate(tdna_st=as.numeric(as.character(tdna_st))) %>% mutate(tdna_end=as.numeric(as.character(tdna_end))) %>% mutate(length_break=abs(tdna_end - tdna_st)) %>% mutate(breakpoint_end=Breakpoint+length_break) %>% mutate(name=paste(Chr,Breakpoint,vector,sample,sep="-"))  %>% select(name,Chr,Breakpoint,breakpoint_end,vector)
write.table(sub_table,file=paste0(path,"annot2.tsv"),sep="\t",row.names = FALSE)
chromoMap(chr,paste0(path,"annot2.tsv"))
```

## Trimming parameters

```{r trimmomatic}
dat<-read.table(text=gsub(" ","\t",readLines(trimmo)),  header=TRUE)
DT::datatable(dat)
```

## Mapping parameters

### Flagstat results

```{r flagstat}
library(tidyverse)
file <- flagstat
if(file.exists(file)){
    flagstat <- read.table(file, sep="\t", header=FALSE) %>% 
      dplyr::mutate(mapped=V3-V4, unmapped=V2-V3, supplementary=V4, 
                    mapped.ratio=round(mapped/(mapped+unmapped)*100,2), 
                    unmapped.ratio=round(unmapped/(mapped+unmapped)*100,2),
                    supplementary.ratio=round(supplementary/(mapped+unmapped)*100,2)) %>% dplyr::select(-V2, -V3, -V4)
    names(flagstat) <- c("sample", "mapped", "unmapped", "supplementary", "mapped.ratio", "unmapped.ratio", "supplementary.ratio")

  }

DT::datatable(flagstat, options = list(rownames = TRUE, pageLength = 10, scrollX = T, dom = 'tip'),class = 'cell-border stripe')
#count
nbr.reads.melt <- flagstat %>% dplyr::select(sample, mapped, unmapped, supplementary) %>% reshape2::melt(id = "sample", value.name = "nbr.reads")
summary.nbr <- nbr.reads.melt %>% group_by(variable) %>% summarise(median=median(nbr.reads), max=max(nbr.reads), min=min(nbr.reads))
nbr.reads.melt <- left_join(nbr.reads.melt, summary.nbr, by = "variable") %>% mutate(x.label=paste0(variable, "\nmedian=", median, "\nmax=", max, "\nmin=", min))
ggplot(data = nbr.reads.melt, aes(y=nbr.reads, x=x.label)) + geom_boxplot()  + xlab("") + ylab("number of reads") + ggtitle("Boxplot representing the number of mapped/supplementary/unmapped reads")+theme_bw()
#ratio
ratio.reads.melt <- flagstat %>% dplyr::select(sample, mapped.ratio, supplementary.ratio) 
names(ratio.reads.melt) <- c("sample", "mapped", "supplementary")
ratio.reads.melt <- ratio.reads.melt %>% reshape2::melt(id = "sample", value.name = "ratio")
summary.ratio <- ratio.reads.melt %>% group_by(variable) %>% summarise( median=round(median(ratio),2), max=round(max(ratio),2), min=round(min(ratio),2))
ratio.reads.melt <- left_join(ratio.reads.melt, summary.ratio, by = "variable") %>% mutate(x.label=paste0(variable, "\nmedian=", median, "%\nmax=", max, "%\nmin=", min, "%"))
ggplot(data = ratio.reads.melt) + geom_boxplot(aes(y=ratio, x=x.label)) + xlab("") + ylab("% reads") + scale_y_continuous(breaks = seq(0, max(ratio.reads.melt$ratio)+10, 20)) + ggtitle("Boxplot representing the percentage of mapped reads and supplementary alignments")+theme_bw()

```

### Coverage analysis
```{r coverage}
cov.list <- lapply(sample_sheet$sample, function(x){
  file <- paste0(workspace,"/03_mapping/coverage/", x, ".coverage")
  if(file.exists(file)){

    read.table(file, sep="\t") %>%  dplyr::filter(V4 > 1)  %>% dplyr::mutate(sample = x)
  }
})

cov <- purrr::reduce(cov.list, rbind) %>% dplyr::mutate(nbr_reads=dplyr::if_else(V4 >= 15, ">15X", "<15X"))
colnames(cov) <- c("rname",	"startpos",	"endpos",	"numreads",	"covbases",	"coverage",	"meandepth",	"meanbaseq",	"meanmapq", "sample", "nbr_reads")
#cov <- cov %>% left_join(bed.tab, by=c("rname","startpos"="start","endpos"="stop")) 
cov.summary <- cov %>% group_by(rname) %>% summarise(min=min(numreads), max=max(numreads), median=round(median(numreads)))
cov <- left_join(cov, cov.summary, by = c("rname") )%>% mutate(x.label=paste0(rname, "\nmedian=", median, "\nmax=", max, "\nmin=", min)) 
for (i in unique(sample_sheet$sample)){
cat("####",i,"\n")
g<-ggplot(filter(cov,sample==i)) + geom_boxplot(aes(y=numreads, x=x.label)) + xlab("") + ylab("#reads") + ggtitle("Plot coverage")+facet_wrap(. ~ sample,scales = "free")+theme(axis.text.x=element_text(angle=90,hjust=1))+theme_bw()
print(g)
cat('\n\n')
}
```

#### Depth analysis
```{r depth_analysis}
cov.list <- lapply(sample_sheet$sample, function(x){

  file <- paste0(workspace,"/03_mapping/depth/", x, ".depth")
  if(file.exists(file)){
    read.table(file, sep="\t")  %>% dplyr::mutate(plant = x)
  }
})

depth <- purrr::reduce(cov.list, rbind)
colnames(depth) <- c("rname", "position", "depth", "plant")
#depth <- depth %>% left_join(bed.tab, by="rname")  %>% filter(position >=start, position <= stop)
heading<-as.character(unique(sample_sheet$sample))
for (i in unique(cov$rname)){
 cat("####",i,"\n")
  p <- ggplot(data = filter(depth,rname== i)) + geom_line(aes(x=position, y=depth, color=plant))   + theme(legend.position="none", axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + ylab("Depth of coverage")+facet_grid(~rname, scales = "free") 
  print(p)
cat('\n\n')
}

```

```{r proportion_of, results='asis',echo=FALSE, eval=FALSE, warning=FALSE, error=FALSE,results=FALSE}
for (i in unique(bed.tab$target)){
 cat("####",i,"\n")
tmp <- cov  %>% filter(target== i) %>% group_by(nbr_reads, target, amplicon) %>% count(name = "#well") 
g1<-ggplot(data = tmp) + geom_bar(aes(x=amplicon, y=`#well`, fill=nbr_reads), stat = "identity") + xlab("") +facet_grid(~target, scales = "free")
g2<-cov %>% filter(target== i) %>% group_by(well, nbr_reads,target) %>% count(name = "amplicon") %>% filter(nbr_reads == ">15X") %>% group_by(amplicon,target) %>% count(name = "#well") %>% ggplot(aes(x = amplicon, y = `#well`)) + geom_bar(stat = "identity") + xlab("nbr of amplicon covered at >15X")
print(g1)
print(g2)
cat('\n\n')
}
```

## Study of SNV and Indels calling

```{r define_r_fct_multiallele, echo=FALSE}
vcfR_split_multiAlt <- function(vcfR.tid, nbrAlt=NULL){

  vcfR.tid.tmp <- 
    dplyr::mutate(vcfR.tid, nb.allele=unlist(lapply(str_split(string = vcfR.tid$gt_AD, pattern = ","), function(x){return(length(x))}))) %>%
    dplyr::mutate(nb.allele=nb.allele-1)

  vcfR.tid.list <- lapply(unique(vcfR.tid.tmp$nb.allele), function(x){

    tmp <- vcfR.tid.tmp %>% dplyr::filter(nb.allele == x) %>%
      tidyr::separate(col = gt_AD, into = c("gt_ref_AD", paste0("gt_alt_AD",1:x)), sep = ",", remove = FALSE) %>%
      tidyr::separate(col = ALT,   into = paste0("ALT" ,1:x), sep = ",", remove = FALSE) %>%
      tidyr::separate(col = AC,    into = paste0("AC" ,1:x),  sep = ",", remove = FALSE) %>%
      tidyr::separate(col = AF,    into = paste0("AF" ,1:x),  sep = ",", remove = FALSE) %>%
      tidyr::separate(col = MLEAF, into = paste0("MLEAF" ,1:x),  sep = ",", remove = FALSE) %>%
      tidyr::separate(col = MLEAC, into = paste0("MLEAC" ,1:x),  sep = ",", remove = FALSE)

    tmp.list <- lapply(1:x, FUN = function(i){

        tmp.split <- dplyr::mutate(tmp,  gt_alt_AD = as.numeric(get(paste0("gt_alt_AD", i))) ,
                                         gt_AR     = as.numeric(get(paste0("gt_alt_AD", i)))/gt_DP, 
                                         ID_old    = ID,
                                         gt_AD_old = gt_AD, gt_AD = paste0(gt_ref_AD, ",", get(paste0("gt_alt_AD", i))),
                                         ALT_old   = ALT,   ALT   = get(paste0("ALT", i)),
                                         AC_old    = AC,    AC    = get(paste0("AC", i)),
                                         AF_old    = AF,    AF    = get(paste0("AF", i)),
                                         MLEAF_old = MLEAF, MLEAF = get(paste0("MLEAF", i)),
                                         MLEAC_old = MLEAC, MLEAC = get(paste0("MLEAC", i)),
                                         ID = paste(CHROM, POS, REF, ALT, sep="_"),
                                   ) %>%
        dplyr::select(-gt_ref_AD, 
                      -paste0("ALT", 1:x), 
                      -paste0("gt_alt_AD", 1:x),  
                      -paste0("AC", 1:x),
                      -paste0("AF", 1:x),
                      -paste0("MLEAF", 1:x),
                      -paste0("MLEAC", 1:x)) %>%
        dplyr::filter(ALT != "*")

        })

        return(tmp.list %>% purrr::reduce(rbind))
    })

    return(purrr::reduce(vcfR.tid.list, rbind))
}
```

```{r , echo=FALSE, warning=FALSE, error=FALSE}
vcfR.inst <- vcfR::read.vcfR(haplotype_caller)

vcfR.tid.list <- vcfR::vcfR2tidy(vcfR.inst)

vcfR.tid <- full_join(vcfR.tid.list$fix, vcfR.tid.list$gt, by = c("ChromKey", "POS")) %>% 
            dplyr::mutate(ID = dplyr::if_else(is.na(ID), paste(CHROM, POS, REF, ALT, sep="_"), ID))

variants <- vcfR_split_multiAlt(vcfR.tid) %>%
  dplyr::mutate(gt_AR_corr = dplyr::if_else(gt_AR < AR.min, 0, gt_AR),
         gt_AR_corr = dplyr::if_else(gt_DP < 4,   0, gt_AR_corr),
         gt_DP_corr = dplyr::if_else(gt_DP < 4,   0, as.numeric(gt_DP)),
         var_type   = dplyr::if_else(str_length(REF) == str_length(ALT), "SNV", dplyr::if_else(str_length(REF) > str_length(ALT), "DEL", "INS")),
         var_length = str_length(ALT) - str_length(REF),
         var_class  = dplyr::if_else(var_length == 0, var_type, 
                      dplyr::if_else(abs(var_length) == 1,  paste0(var_type, "1"),
                      dplyr::if_else(abs(var_length) >  50, paste0(var_type, ">50"), paste0(var_type, "<50")))))

variants.control <- variants %>% filter(Indiv %in% dplyr::filter(sample_sheet, condition == "WT")$plant, gt_AR_corr > 0) 
variants.freq_corr <- variants %>% filter(gt_AR_corr > 0, !is.na(gt_AR_corr)) %>% 
  select(ID, Indiv, target,amplicon) %>% unique() %>% group_by(ID, target,amplicon) %>% count(name="freq")

variants <- variants %>% dplyr::filter(!ID %in% variants.control$ID) %>% dplyr::full_join(variants.freq_corr, by = c("ID", "target","amplicon"))
#DT::datatable(variants)
```