### 2023-04-09 script/tools.Rmd

```{r}

config_advanced_file<-paste0(workspace,"/00_logs/config_advanced")
config_ad<-read_yaml(config_advanced_file)
```
# Tools used

```{r tools_used}
space=paste0(workspace,"/00_logs/envs")
for (file in list.files(space)){
  env<-read_yaml(paste0(space,"/",file))
  for (envi in env$dependencies){
    line=sub("=="," version ",envi)
    cat(line,"\n")
  }
}
```

# Trimming
trimmomatic PE `r config_ad$params_trimmomatic ` input_r1 input_r2 output_r1 output_r2 TRAILING:3 2> log

# Alignments

## alignments
bwa mem -R "@RG\tID:sample\tSM:sample\tLB:Illumina\tPL:Illumina\tPU:sample" -M `r config_ad$params_bwa` reference input_fq1 input_fq2 > output

## convert sam to bam
picard SortSam --INPUT input --OUTPUT output --SORT_ORDER coordinate `r config_ad$params_sort_sam`

## mark duplicates
picard MarkDuplicates --INPUT input --OUTPUT{output -M metrics `r config_ad$params_mark_duplicate`

# QC bam

## flagstat
samtools flagstat `r config_ad$params_flagstat ` input > output

## samtools coverage
samtools coverage -o output `r config_ad$params_samtools_coverage ` input

## samtools depth
samtools depth `r config_ad$params_samtools_depth`  input > output

# SV Calling

## Run haplotyCallze
gatk HaplotypeCaller -R reference -I input -O output -L region --native-pair-hmm-threads 16 `r config_ad$params_haplotype_caller`

## Run GenomicsDB
gatk GenomicsDBImport variant --genomicsdb-workspace-path output --intervals {params.region} `r config_ad$params_genomics_db`

## Run GenotypeGVCF
gatk GenotypeGVCFs --reference reference -V gendb://input -O output `r config_ad$params_genotype_gvcf`

## Merge gvcf
picard GatherVcfs list_of_vcf_files O=output `r config_ad$params_gathervcf`

## Build snpeff database
snpEff build -c output `r config_ad$params_snpeff_build` -gff3 -v reference -dataDir .

## Annotate variant
snpEff eff -c input `r config_ad$params_snpeff_ann` -dataDir . reference -s output.html input > output

# Run pindel
## run pindel
pindel `r config_ad$params_run_pindel`  -f reference -i input -o output

## convert pindel to vcf
pindel2vcf `r config_ad$params_convert_pindel` -r reference -P sample name -R xxx -d xx -v output

## Index vcf files and merge them
tabix input
bcftools merge -m none -Oz `r config_ad$params_merge_vcf_files` -o output list_vcf_files

# Run tdnascan for vaector check

## Cut multi-fasta vector file in individual files
faSplit byname input output_directory `r config_ad$params_cut_vector`

## run tdnascan
python2.7 tdnascan.py -1 input_fq1 -2 input_fq2 -t vector_sequence -g reference -p vector_name -i script_directory -d output_directory `r config_ad$params_tdnascan`

## run tdnascan annotate
python2.7 tdnascan_annotate.py -i input -f annotation_file -o output `r config_ad$params_tdnascan_annotate`