### 2023-04-11 script/quality_control.Rmd
add rmd for quality control

## Trimming parameters

```{r trimmomatic}
trimmo<-paste0(workspace,"/01_sequence_qc/log/trimmomatic.log")
dat<-read.table(text=gsub(" ","\t",readLines(trimmo)),  header=TRUE)
DT::datatable(dat)
```

## HTML

## Mapping parameters

### Flagstat results

```{r }

file <- paste0(workspace,"/02_mapping/flagstat/concatenate_flagstat.txt")
if(file.exists(file)){
    flagstat <- read.table(file, sep="\t", header=FALSE) %>% 
      dplyr::mutate(mapped=V3-V4, unmapped=V2-V3, supplementary=V4, 
                    mapped.ratio=round(mapped/(mapped+unmapped)*100,2), 
                    unmapped.ratio=round(unmapped/(mapped+unmapped)*100,2),
                    supplementary.ratio=round(supplementary/(mapped+unmapped)*100,2)) %>% dplyr::select(-V2, -V3, -V4)
    names(flagstat) <- c("sample", "mapped", "unmapped", "supplementary", "mapped.ratio", "unmapped.ratio", "supplementary.ratio")

  }

DT::datatable(flagstat, options = list(rownames = TRUE, pageLength = 10, scrollX = T, dom = 'tip'),class = 'cell-border stripe')

#count
nbr.reads.melt <- flagstat %>% dplyr::select(sample, mapped, unmapped, supplementary) %>% reshape2::melt(id = "sample", value.name = "nbr.reads")

p <- ggplot(data = nbr.reads.melt, aes(y=nbr.reads, x=variable)) + geom_boxplot()  + 
  xlab("") + ylab("number of reads") + ggtitle("Boxplot representing the number of mapped/supplementary/unmapped reads")+theme_bw()

print(p)

#ratio
ratio.reads.melt <- flagstat %>% dplyr::select(sample, mapped.ratio, supplementary.ratio) 
names(ratio.reads.melt) <- c("sample", "mapped", "supplementary")

ratio.reads.melt <- ratio.reads.melt %>% reshape2::melt(id = "sample", value.name = "ratio")

summary.ratio <- ratio.reads.melt %>% group_by(variable) %>% 
  summarise( median=round(median(ratio),2), max=round(max(ratio),2), min=round(min(ratio),2))

p <- ggplot(data = ratio.reads.melt) + geom_boxplot(aes(y=ratio, x=variable)) + 
  xlab("") + ylab("% reads") + scale_y_continuous(breaks = seq(0, max(ratio.reads.melt$ratio)+10, 20)) + 
  ggtitle("Boxplot representing the percentage of mapped reads and supplementary alignments")+theme_bw()

print(p)

```

### Coverage analysis
```{r coverage}
cov.list <- lapply(sample_sheet$sample, function(x){
  file <- paste0(workspace,"/02_mapping/coverage/", x, ".coverage")
  if(file.exists(file)){

    read.table(file, sep="\t") %>%  dplyr::filter(V4 > 1)  %>% dplyr::mutate(sample = x)
  }
})

cov <- purrr::reduce(cov.list, rbind) %>% dplyr::mutate(nbr_reads=dplyr::if_else(V4 >= 15, ">15X", "<15X"))
colnames(cov) <- c("rname",	"startpos",	"endpos",	"numreads",	"covbases",	"coverage",	"meandepth",	"meanbaseq",	"meanmapq", "sample", "nbr_reads")
#cov <- cov %>% left_join(bed.tab, by=c("rname","startpos"="start","endpos"="stop")) 
#cov.summary <- cov %>% group_by(rname) %>% summarise(min=min(numreads), max=max(numreads), median=round(median(numreads)))
#cov <- left_join(cov, cov.summary, by = c("rname") )%>% mutate(x.label=paste0(rname, "\nmedian=", median, "\nmax=", max, "\nmin=", min))

p <-ggplot(data = cov) + geom_boxplot(aes(y=meandepth, x=rname)) + 
  xlab("") + ggtitle("Plot coverage")  + 
  theme(axis.text.x=element_text(angle=90,hjust=1))+theme_bw()

print(p)

```