### 2023-04-11 script/report.Rmd

---
title: "Report"
author: " "
date: "2022-08-16"
output:
    html_document:
        highlight: tango
        number_sections: no
        theme: default
        toc: yes
        toc_depth: 3
        toc_float:
            collapsed: no
            smooth_scroll: yes
params:
  DP.min: DP.min
  AR.min: AR.min
  result_dir: result_dir
  rmd: "report.Rmd"
  AR: AR
---

```{r setup, include=FALSE}
# title: "`r params$title`"
workspace <- "/Users/nbentahar/Documents/PROJECTs/DEV/msgenova/examples/soj/results/"
#workspace<-params$result_dir

sample_sheet.file <- "/Users/nbentahar/Documents/PROJECTs/DEV/msgenova/examples/soj/sample_sheet_soj_3samples.txt"

knitr::opts_chunk$set(root.dir=workspace, echo = FALSE , warning=FALSE, error=TRUE, message=FALSE)
```

```{r }
library(DT)
library(ggplot2)
library(vcfR)
library(dplyr)
library(yaml)

```

```{r}

sample_sheet <- read.table(file = sample_sheet.file, header = TRUE, sep = "\t")

```

```{r quality_control.Rmd}

  out <- NULL

  out <- c(out, knitr::knit_child("quality_control.Rmd"), quiet=TRUE)

knitr::asis_output(out)

```

```{r SNV_InDels_calling.Rmd, eval=FALSE}

vcf.file <- "/Users/nbentahar/Documents/PROJECTs/DEV/msgenova/examples/soj/results/03_snv_indels_calling/SOJ_snv_indel.vcf"

  out <- NULL

  out <- c(out, knitr::knit_child("SNV_InDels_calling.Rmd"), quiet=TRUE)

knitr::asis_output(out)

```

```{r SV_calling.Rmd, eval=FALSE}

  out <- NULL

  out <- c(out, knitr::knit_child("SV_calling.Rmd"), quiet=TRUE)

knitr::asis_output(out)

```

```{r tDNA_scan.Rmd, eval=FALSE}

  out <- NULL

  out <- c(out, knitr::knit_child("tDNA_scan.Rmd"), quiet=TRUE)

knitr::asis_output(out)

```

```{r tools, eval=FALSE}

  out <- NULL

  out <- c(out, knitr::knit_child("tools.Rmd"), quiet=TRUE)

knitr::asis_output(out)

```