### 2023-04-14 script/per_sample_ins_desc.Rmd
Add tdnascan report.

### `r sample` {.tabset}

#### Summary of insertion reads

```{r insertion, echo=FALSE}

insRead.list <- list()

for(vector_i in perSample_vectorList[[sample]]$both.list){

  insertion_read_file <- paste0(path_results,"/results/05_tdnascan","/",sample,"/",vector_i,"/4.insertionRead.txt")

  if(file.exists(insertion_read_file) & file.size(insertion_read_file) != 0 ){

    insRead.list[[paste0(sample,"_",vector_i)]] <- read.table(insertion_read_file,sep="\t") %>% 
      dplyr::mutate(sample = sample) %>% 
      dplyr::mutate(vector = vector_i)

  }
}

insRead <- purrr::reduce(insRead.list, rbind)

insRead.summary <- insRead %>% 
dplyr::mutate(insRead,V1=str_remove_all(V1,pattern="[\\(\\)]")) %>%   tidyr::separate(V1,into=c("chromosome","position","type","tdna_border","tdna_position","orientation"),sep=",") %>% dplyr::mutate(position=as.numeric(position)) %>% dplyr::group_by(chromosome,position,type,tdna_border,tdna_position,orientation,vector) %>%
dplyr::count()

DT::datatable(insRead.summary)
```

```{r depthNGS, eval=TRUE, echo=FALSE}

depth <- read.table(paste0(path_results,"/results/02_mapping/depth/",sample,".depth"),h=TRUE) 
colnames(depth) <- c("rname", "position", "depth")
SIZE=10000

p <- list()

for(vector_j in unique(ins$vector)){

  ins.sample <- dplyr::filter(ins, sample == !!sample, vector == !!vector_j) %>% 
    dplyr::group_by(vector, insertion_id)

  depth.filt.sample <- list()

  if(dim(ins.sample)[1] != 0){

    for(i in 1:dim(ins.sample)[1]){

      depth.filt <- dplyr::filter(depth, 
                                  rname == ins.sample[i,]$Chromosome, 
                                  position >= (ins.sample[i,]$Position-SIZE),
                                  position <= (ins.sample[i,]$Position+SIZE))

      depth.filt <- dplyr::mutate(depth.filt, 
      insertion_id = paste0(sample,"_",ins.sample[i,]$vector,"_",rname,"_ins_",i), vector=ins.sample[i,]$vector)

      depth.filt.sample[[i]] <- depth.filt[seq(from=1,to=dim(depth.filt)[1],by=20),]

    }
    depth.plot <- purrr::reduce(depth.filt.sample, rbind)

    p[[vector_j]] <- ggplot(depth.plot) + geom_bar(aes(x=position, y=depth),stat="identity",colour="blue") + 
      theme(legend.position="none", axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
      ylab("Depth of coverage") + facet_wrap(vars(insertion_id),scales = "free", ncol=1) + 
      geom_vline(data=ins.sample, aes(xintercept = Position ,col="red")) + 
      geom_text(data=ins.sample,aes(x=Position,y=4,label=paste0("ins_pos: ",Position),colour="red",fontface = "bold"))
    #print(p[[vector_j]])
  }
}

```

#### Depth along insertion sites +- 10 kb {.tabset}

```{r, eval=TRUE, echo=FALSE, results='asis', fig.height=15}
for(vector_j in unique(ins$vector)){
  cat("##### Vector", vector_j, "\n")
  print(p[[vector_j]])
  cat("\n\n")
}
```