### 2023-04-14 script/per_sample_sv_desc.Rmd

### `r sample` {.tabset}

#### summary of homozygous SV

```{r sv_spec_plot_hom, echo=FALSE}

bam <- paste0(sample,".bam")

SV.NGS.hom <-  SV %>%  dplyr::filter(
    .data[[bam]]  %in% c("1/1"),SPEC == 1) %>% dplyr::arrange(.,CHROM,POS)

if(dim(SV.NGS.hom)[1] >=1){
SV.NGS.hom  %>% group_by(CHROM,SVTYPE) %>% count() %>% ggplot() + geom_bar(aes(x=CHROM,y=n,fill=SVTYPE),stat="identity") + scale_fill_manual(values=c("lightsalmon","royalblue1","green","magenta1","yellow"),drop=FALSE) + xlab("Chromosoms")
}

```

#### table of homozygous SV

```{r sv_spec_plot_table_hom, echo=FALSE}
DT::datatable(SV.NGS.hom %>% dplyr::select(.,-contains(".bam")))
```

#### summary of heterozygous SV

```{r sv_spec_plot_het, echo=FALSE}
bam <- paste0(sample,".bam")

SV.NGS.het <-  SV %>%  dplyr::filter(
    .data[[bam]]  %in% c("0/1"),SPEC == 1) %>% dplyr::arrange(.,CHROM,POS)

if(dim(SV.NGS.het)[1] >=1){
SV.NGS.het %>% group_by(CHROM,SVTYPE) %>% count() %>% ggplot() + geom_bar(aes(x=CHROM,y=n,fill=SVTYPE),stat="identity") + scale_fill_manual(values=c("lightsalmon","royalblue1","green","magenta1","yellow"),drop=FALSE)
}
```

#### table of heterozygous SV

```{r sv_spec_plot_table_het, echo=FALSE}
DT::datatable(SV.NGS.het %>% dplyr::select(.,-contains(".bam")))
```