### 2023-04-14 script/per_sample_vector_depth.Rmd

### `r sample`

```{r depthVec, eval=TRUE, echo=FALSE}

file <- paste0(path_results,"/results/02_mapping/depth/",sample,".depth")

if(file.exists(file)){
  depth <- read.table(file,h=TRUE) 
}

colnames(depth) <- c("rname", "position", "depth")

depth.vector <- dplyr::filter(depth, rname %in%  perSample_vectorList[[sample]]$both.list)

if(dim(depth.vector)[1] > 0){

ggplot(depth.vector) + geom_bar(aes(x=position, y=depth),stat="identity",colour="blue") + 
  theme(legend.position="none", axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  ylab("Depth of coverage") + facet_wrap(vars(rname),scales = "fixed",ncol=1)
}

```