### 2023-04-14 script/tDNA_scan.Rmd

# parameters

```{r param}
path_results <- workspace
samplesheet <- sample_sheet
vector.sep <- ","
nb_common_samples=10
SVsize_CutOff = 500
```

# test sample sheet

```{r testSS, eval=TRUE, echo=FALSE}

# test de l'existence de sample_sheet:
# !! des colonnes additionnelles sont attendues pour le sample_sheet !!
# Si celles-ci sont complétées au démarrage du pipeline, cela posera
# problème avec le pipeline snakemake
# tdna (required): 1 to many separate by ","
# vector (optional): NA to many separate by ","

if(dir.exists(path_results) & file.exists(samplesheet)){
  sample_sheet <- read.table(samplesheet,sep="\t",header=TRUE)
}

# construction of the list of searched tdna and/or vector

perSample_vectorList <- as.list(1:dim(sample_sheet)[1])

for(k in 1:dim(sample_sheet)[1]){
  tmp.l <- list()
  tmp.l[["tdna.list"]] <- stringr::str_split(sample_sheet[k,"tdna"], pattern = vector.sep)[[1]]
  tmp.l[["vector.list"]] <- stringr::str_split(sample_sheet[k,"vector"],pattern = vector.sep)[[1]]
  tmp.l[["both.list"]] <- na.omit(c(tmp.l[["tdna.list"]],tmp.l[["vector.list"]]))
  perSample_vectorList[[k]] <- tmp.l
}
names(perSample_vectorList) <- sample_sheet$sample

```

# Load libraries

```{r load.lib, echo=FALSE,message=FALSE}
library(tidyverse)
library(vcfR)
```

# tDNA coverage statistics in comparison with other chromosoms:

This statistics will indicate the number of times the tDNA may be inserted and if it
is completely inserted.

## Mean depth:

```{r depth, echo=FALSE, warning=FALSE}

# reads the statistics of chromosoms and vectors mean depth

cov.list <- lapply(sample_sheet$sample, function(x){

  file <- paste0(path_results,"/results/02_mapping/coverage","/", x, ".coverage")

  if(file.exists(file)){
    read.table(file, sep="\t") %>%  dplyr::filter(V4 > 1)  %>% dplyr::mutate(sample = x)
  }
})

if(length(cov.list) != 0){

  cov <- purrr::reduce(cov.list, rbind)

  colnames(cov) <- c("rname",	"startpos",	"endpos",	"numreads",	
                     "covbases",	"coverage",	"meandepth",	"meanbaseq",	"meanmapq", "sample")

  # Mean depth (log10)
  g1 <- ggplot(cov) + geom_bar(aes(y=meandepth, x=rname),stat = "identity") + 
    xlab("") + ylab("Mean Depth") + ggtitle("Mean depth (log10)") + facet_wrap( ~ sample) +
    theme(axis.text.x=element_text(angle=90,hjust=1)) + theme_bw() + scale_y_continuous(trans='log10',limits=c(1,NA)) +
    coord_flip() + geom_text(aes(x = rname,y=10,label=paste0(round(meandepth,0)," x")),color="red")
  g1
}
```

## Chromosoms coverage (%):

```{r coverage, echo=FALSE}
if(length(cov) != 0){
  g2<-ggplot(cov) + geom_bar(aes(y=coverage, x=rname),stat = "identity") + xlab("") + ylab("% coverage") + ggtitle("Chromosoms coverage (%)")+facet_wrap( ~ sample)+theme(axis.text.x=element_text(angle=90,hjust=1)) + theme_bw() + coord_flip()
  g2
}
```

## Depth of coverage along vectors:

<!-- ### Graph per sample -->

```{r , message=FALSE, warning=FALSE, echo=FALSE}

  out <- NULL

  for (sample  in  unique(sample_sheet$sample)) {
      out <- c(out, knitr::knit_child("per_sample_vector_depth.Rmd", quiet=TRUE))  
    }

knitr::asis_output(out)
```

# TDNAscan results:

tdnascan was used to detect tdna insertion [@sun_tdnascan_2019].

## Table of detected insertion: {.tabset}

* Results description:
  + Chromosome: Chromosome number
  + Position: Start position of insertion in the reference genome
  + \#soft.clipped: The soft-clipped reads (CLR) are reads where one partial read of a single read perfectly mapped to the plant reference genome and the other partial read of the same single read perfectly mapped to the inserted T-DNA. 
  + \#Discordant: In tdnascan, the discordant reads (DIR) are defined as one read of a pair successfully mapped to the plant reference genome and the other read of the same pair mapped to part of the inserted T-DNA.
  + tdna.start: Start position of T-DNA sequence
  + tdna.end: End position of T-DNA sequence
  + orientation: Forward or reverse T-DNA inserted to reference genome
  + insertion zygotie: #soft.clipped/(\#soft.clipped + \#span) with \#span = \# reads MAPQ > 30 which span the Insertion site with at least 5bp at each side
  + annotation:
  + sample: 
  + insertion_id:

### Overview of exact insertion's positions

```{r bed.tab.plot, echo =FALSE,message=FALSE}

# Concatenate all tdnascan results

ins.list <- list()

for(x in sample_sheet$sample){

  for(vector_i in perSample_vectorList[[x]]$both.list){

    file <- list.files(path=paste0(path_results,"/results/05_tdnascan","/",x,"/",vector_i,"/"), 
                       pattern="annotated.bed", full.names = TRUE)

    if(length(file) != 0){
      if(file.size(file) != 0){

        tmp <- read.table(file, sep="\t",h=FALSE) %>% dplyr::mutate(sample = x, vector = vector_i)

        if(dim(tmp)[1] != 0){

          ins.list[[paste0(x,"_",vector_i)]] <- 
            dplyr::mutate(tmp, insertion_id = paste0(x,"_",vector_i,"_",V1,"_ins_",1:dim(tmp)[1]))

        }
      }
    }
  }
}

ins <- purrr::reduce(ins.list, rbind) 
ins <- separate(ins,V3,into=c("CLR","DIR"),sep=",")
ins <- separate(ins,V4,into=c("tdna_st","tdna_end"),sep=",")
ins <- mutate(ins,CLR = str_extract(CLR,pattern="(\\d)+"),
              DIR=str_extract(DIR,pattern="(\\d)+"),
              tdna_st=str_extract(tdna_st,pattern="(\\d)+"),
              tdna_end=str_extract(tdna_end,pattern="(\\d)+"))

names(ins)<-c("Chromosome","Position","#soft.clipped","#Discordant",
              "tdna.start","tdna.end","orientation","insertion zygotie",
              "annotation","sample","vector","insertion_id")

Chrom <- dplyr::filter(cov,rname %in% ins$Chromosome) %>% dplyr::select(.,c("rname","startpos","endpos")) 
Chrom <- dplyr::distinct(Chrom) %>% dplyr::rename(.,Chromosome=rname)

 ggplot(data=ins)+
     geom_rect(data=Chrom, mapping = aes(xmin=startpos,xmax=endpos,ymin=0.2,ymax=0.3),color="grey")+
     facet_grid(sample~Chromosome,space="free_x",scales="free_x") + 
   geom_jitter(data=ins, aes(x=Position,y=1,color=vector),size=2,shape=25) + ylim(c(0,2)) + 
    theme(axis.text.y=element_blank(),axis.ticks.y =element_blank(),axis.text.x= element_text(angle = 90), axis.title.y=element_blank(),axis.title.x=element_blank(),strip.background = element_blank(),panel.grid.minor = element_blank(),strip.text.y = element_text(angle=-90),panel.grid.major = element_blank(),strip.placement = "outside")

```

### Exact insertion's positions table

```{r bed.tab.view, echo =FALSE,message=FALSE}
DT::datatable(ins)
```

### Overview of approximativ breakpoint's positions

Overview of detected insertion before filtering for nearby breakpoints:

```{r nofiltre_view, echo = FALSE, message = FALSE}

ins.list.raw <- list()

for(x in sample_sheet$sample){

  for(vector_i in perSample_vectorList[[x]]$both.list){

    file <- list.files(paste0(path_results,"/results/05_tdnascan","/",x,"/",vector_i,"/"),
                       pattern="insertion.bed", full.names = TRUE)

    if(length(file) != 0){

      ins.list.raw[[paste0(x,"_",vector_i)]] <- read.table(file, sep="\t",h=TRUE) %>% 
        dplyr::mutate(sample = x) %>% dplyr::mutate(vector = vector_i) 
    }
  }
}

ins_raw <- purrr::reduce(ins.list.raw, rbind)

if(length(ins_raw) > 0){

  ins_raw <- tidyr::separate(ins_raw,SuppRead,into=c("CLR","DIR"),sep=",")
  ins_raw <- tidyr::separate(ins_raw,TDNA_info,into=c("tdna_st","tdna_end"),sep=",")
  ins_raw <- dplyr::mutate(ins_raw,CLR=str_extract(CLR,pattern="(\\d)+"),
                           DIR=str_extract(DIR,pattern="(\\d)+"),
                           tdna_st=str_extract(tdna_st,pattern="(\\d)+"),
                           tdna_end=str_extract(tdna_end,pattern="(\\d)+"))
  ins_raw <- dplyr::filter(ins_raw,str_detect(Breakpoint,"~")==TRUE)
  ins_raw <- dplyr::mutate(ins_raw,Breakpoint=as.numeric(str_remove(Breakpoint,"~")))

  names(ins_raw)<-c("Chromosome","Breakpoint","#soft.clipped","#Discordant",
                    "tdna.start","tdna.end","orientation","insertion zygotie",
                    "sample","vector")

  Chrom <- dplyr::filter(cov,rname %in% ins_raw$Chromosome) %>% dplyr::select(.,c("rname","startpos","endpos")) 
  Chrom <- dplyr::distinct(Chrom) %>% dplyr::rename(.,Chromosome=rname)

  ggplot(data=ins_raw)+
    geom_rect(data=Chrom, mapping = aes(xmin=startpos,xmax=endpos,ymin=0.2,ymax=0.3),color="grey")+
    facet_grid(sample~Chromosome,space="free_x",scales="free_x") + 
    geom_jitter(data=ins_raw, aes(x=Breakpoint,y=1,color=vector),size=2,shape=25) + ylim(c(0,2)) + 
    theme(axis.text.y=element_blank(),axis.ticks.y =element_blank(),axis.text.x= element_text(angle = 90), axis.title.y=element_blank(),axis.title.x=element_blank(),strip.background = element_blank(),panel.grid.minor = element_blank(),strip.text.y = element_text(angle=-90),panel.grid.major = element_blank(),strip.placement = "outside", ) 
}

```

### Approximativ breakpoints table

Table of detected insertion before filtering for nearby breakpoints:

```{r nofiltre_table, echo =FALSE,message=FALSE}
DT::datatable(ins_raw)
```

## Informativs reads description:

* This table describes informatives reads. The reads which mapped to part of the tDNA and which cover the insertion sites. 
  + for CLR: The reads are retained if they satisfied the criterion of the difference in the number of soft clipped bases in the TDNA and in the genome (<5). See the figure below.     
  + for DIR: The reads are retained if one mate match completly to the tDNA and the other to the genome.
  + TSP: tdna start border 
  + TEP: tdna end boder

<!-- ### Results per sample -->

```{r , message=FALSE, warning=FALSE, echo=FALSE}

  out <- NULL

  for (sample  in  unique(sample_sheet$sample)) {
      out <- c(out, knitr::knit_child("per_sample_ins_desc.Rmd", quiet=TRUE))

    }

knitr::asis_output(out)
```

# Do the tdna-insertion lines have specifics Structural Variation ?

* pindel [ref] was used to detect the following structural variations.
  + INV: Inversion (min_length=50): 
  + DUP:TANDEM: tandem duplication
  + INS: Insertion: Domestic insertion (DINS) of Foreign Insertion (FINS) or Translocation (TRA)
  + DEL: Deletion
  + RPL: Replacement: describes an insertion event around the breakpoint of a deletion event, which could in fact capture DUP, INV, DINS, FINS and TRA, thus occasionally resulting in duplicate SV calls with different assigned SV types.

* NB: 
  + pindel as a poor sensitivity to detect DINS or TRANS. It may not help to confirm tDNAscan results but may help to detect 
other SVs that might be induced by the tDNA lines transformation.
We will include other sv caller results such as delly to cross tdnascan results and to ensure of the lines background.
  + SVs were classified according to their genotypes: 1/1 = homozygous for the alternate allele and 1/0 = heterozygous
  They were not confirm thanks to igv, or by filtering allelic depth. 
  + A filter was applied to keep SV longer than 500.
  + A column NEXT_SV was add to indicate the distance (in base) to the next SV: it can be a common or specific one. 
  It will help to filter sv classified in specific sv whereas they are common to samples.

```{r sv,eval=TRUE, message=FALSE,echo=FALSE}

pindel <- paste0(path_results,"/results/04_pindel/",path_results,"_sv.vcf.gz")

# import pindel vcf

if(file.exists(pindel)){
  vcfR.inst <- vcfR::read.vcfR(pindel,verbose = FALSE)

  nsample <- dim(vcfR.inst@gt)[2]-1

  # transform to tidy
  vcfR.tid.list <- suppressMessages(vcfR::vcfR2tidy(vcfR.inst))

  # filter SV by SVsize_CutOff 
  sv <- vcfR.tid.list[[1]] %>% dplyr::mutate(.,ID=c(1:dim(vcfR.tid.list[[1]])[1]))
  sv.filt <- sv[which(abs(sv$SVLEN) > SVsize_CutOff),]

  if(dim(sv.filt)[1] > 0){
    # filter geno by ID in sv and by GT: 1/1 or 0/1
    geno <- vcfR.tid.list[[2]]  %>%  dplyr::mutate(.,ID=rep(c(1:dim(vcfR.tid.list[[1]])[1]),times=nsample))
    geno.filt <-  dplyr::filter(geno,ID %in% sv.filt$ID, gt_GT %in% c("1/1","0/1")) %>% dplyr::arrange(ID)
  }

  if(dim(geno.filt)[1] > 0){
    geno_tmp <- geno.filt %>% dplyr::select(ID,gt_GT,Indiv,ID) %>% tidyr::pivot_wider(names_from=Indiv,values_from=gt_GT)

    SV <- dplyr::left_join(geno_tmp ,sv.filt,by="ID")  %>% dplyr::arrange(ID) %>% dplyr::mutate(.,SVTYPE=as.factor(SVTYPE)) %>%  dplyr::select(.,ID,CHROM,POS,END,SVLEN,SVTYPE,contains(".bam")) %>% dplyr::mutate(.,DIST_NEXTSV=c(diff(POS),NA))

    tmp <- dplyr::select(SV,contains(".bam"))
    SV$SPEC <- rowSums(apply(! is.na(tmp),2,as.numeric))
  }
}
```

## Overview of SV repartition on chromosoms:

```{r SV_overview, echo=FALSE}
Chrom <- dplyr::filter(cov,rname %in% SV$CHROM) %>% dplyr::select(.,c("rname","startpos","endpos")) 
Chrom <- dplyr::distinct(Chrom) %>% dplyr::rename(.,CHROM=rname)

 ggplot(data=SV)+
     geom_rect(data=Chrom, mapping = aes(xmin=startpos,xmax=endpos,ymin=0,ymax=0.1),color="grey")+
     facet_grid(CHROM ~.,space="free_x",scales="free_x") + 
   geom_jitter(data=SV, aes(x=POS,y=1,color=as.factor(SPEC)),size=2,shape=25) + ylim(c(0,2)) + xlab("Positions") + 
    theme(axis.text.y=element_blank(),axis.ticks.y =element_blank(),axis.text.x= element_text(angle = 90), axis.title.y=element_blank(),axis.title.x=element_blank(),strip.background = element_blank(),panel.grid.minor = element_blank(),strip.text.y = element_text(angle=-90),panel.grid.major = element_blank(),strip.placement = "outside")+
          scale_colour_discrete(name  = "Number of \n samples per SV")

```

## Common SV {.tabset}

### summary of common SV

```{r commonsv_sum,eval=TRUE, message=FALSE,echo=FALSE}
sv_common <- dplyr::filter(SV, SPEC !=1 )
sv_common %>% group_by(CHROM,SVTYPE) %>% count() %>% ggplot() + geom_bar(aes(x=CHROM,y=n,fill=SVTYPE),stat="identity") + scale_fill_manual(values=c("lightsalmon","royalblue1","green","magenta1","yellow"),drop=FALSE) + xlab("Chromosoms")
```

### table

```{r commonsv_tab,eval=TRUE, message=FALSE,echo=FALSE}
DT::datatable(sv_common)
```

## Specific SV per sample:

<!-- ### Results per sample -->

```{r , message=FALSE, warning=FALSE, echo=FALSE}

  out <- NULL

  for (sample  in  unique(sample_sheet$sample)) {
      out <- c(out, knitr::knit_child("per_sample_sv_desc.Rmd", quiet=TRUE))

    }

knitr::asis_output(out)
```