### 2023-07-07 script/report.Rmd

---
title: "Report"
author: " "
date: "2022-08-16"
output:
    html_document:
        highlight: tango
        number_sections: yes
        theme: default
        toc: yes
        toc_depth: 3
        toc_float:
            collapsed: yes
            smooth_scroll: yes
params:
  DP.min: DP.min
  AR.min: AR.min
  qual.min: qual.min
  result_dir: result_dir
  sample_sheet: sample_sheet
  projectName : projectName
  rmd: "report.Rmd"
  AR: AR
---

```{r}
# title: "`r params$title`"
workspace <-params$result_dir
sample_sheet.file <- params$sample_sheet
DP.min <- params$DP.min
AR.min <- params$AR.min
qual.min <- params$qual.min
projectName <- params$projectName
#Genome.file <- "/save/project/ijpb/bank/GENOME/Ppatens-patens-v3.3-WG/Ppatens-patens-v3.3-WG.fa"
#projectName <- "PE-WGS"

#workspace <- "/Users/nbentahar/Desktop/msgenova_test/PE_WGS/results"
#workspace <- "/Users/nbentahar/Documents/PROJECTs/DEV/msgenova/examples/soj/results/"

#sample_sheet.file <- "/Users/nbentahar/Desktop/msgenova_test/PE_WGS/sample_sheet_PE.txt"
#sample_sheet.file <- "/Users/nbentahar/Documents/PROJECTs/DEV/msgenova/examples/soj/sample_sheet_soj_3samples.txt"

# DP.min <- 4
# AR.min <- 0.8
# qual.min <- 30
nbr.alt.min <- 3
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE , warning=FALSE, error=FALSE, message=FALSE, out.width="70%", fig.show="hold", eval = FALSE)
```

```{r}
# title: "`r params$title`"
workspace <-params$result_dir
sample_sheet.file <- params$sample_sheet
DP.min <- params$DP.min
AR.min <- params$AR.min
qual.min <- params$qual.min
projectName <- params$projectName
#Genome.file <- "/save/project/ijpb/bank/GENOME/Ppatens-patens-v3.3-WG/Ppatens-patens-v3.3-WG.fa"
#projectName <- "PE-WGS"

#workspace <- "/Users/nbentahar/Desktop/msgenova_test/PE_WGS/results"
#workspace <- "/Users/nbentahar/Documents/PROJECTs/DEV/msgenova/examples/soj/results/"

#sample_sheet.file <- "/Users/nbentahar/Desktop/msgenova_test/PE_WGS/sample_sheet_PE.txt"
#sample_sheet.file <- "/Users/nbentahar/Documents/PROJECTs/DEV/msgenova/examples/soj/sample_sheet_soj_3samples.txt"

# DP.min <- 4
# AR.min <- 0.8
# qual.min <- 30
nbr.alt.min <- 3
```

```{r}
#libraries
library(DT)
library(ggplot2)
library(vcfR)
library(dplyr)
library(tidyr)
library(stringr)
library(dendextend) 
#library(GenomicRanges)
source("functions.R")
```

# Objective

# Input data
## Genome version
```{r, eval=FALSE}
print(basename(Genome.file))
```

## Samples
```{r}
# error message if sample sheet not exist
if(!file.exists(sample_sheet.file)) stop("Error: ")

# read sample sheet
sample_sheet <- read.csv(file = sample_sheet.file, header = TRUE, sep = "\t")
```

```{r}
sampleNames <- unique(sample_sheet$sample)
```

```
List of samples : `r sampleNames`
```

```{r list samples}

dplyr::select(sample_sheet, sample, condition) %>% unique() %>% 
  dplyr::group_by(condition) %>% dplyr::count(name = "nb_samples") %>% 
  DT::datatable(rownames = FALSE, options = list(rownames = TRUE, pageLength = 6, scrollX = T, dom = 'tip'),
                class = 'cell-border stripe', caption = "List of conditions")

```

# MsGenoVa workflow

![alt text here](workflow.png)